### 2019-04-11 sections/abstract.tex
Imported template stuff

\chapter*{Abstract}
\todo[inline,color=blue!10]{This should be done when most of the rest of the document is finished. Be concise, introduce context, problem, known approaches, your solution, your findings.}