### 2019-07-18 sections/acknowledgements.tex
Acknowledgements

\chapter*{Acknowledgements}
Throughout the writing of this dissertation I have received a great deal of support and assistance. I would first like to thank my supervisor, Dr. A.M. Oprescu , whose expertise was invaluable in the academic writing and mathematical aspects of this thesis.

Next, I would like to thank my peer student Sander Meester for always being there when I was stuck. Whenever I needed to test some ideas you would always provide realistic insights. The long conversations about the problems I have been dealing with in my thesis would not have been solved in the same way if it wasn't for you. Your proofreading efforts helped take my writing to the next level.

Finally, I would like to acknowledge my colleagues from my internship at SIG for their wonderful collaboration. You supported me greatly and were always willing to help me.