### 2019-08-16 sections/experimentalsetup.tex
Thesis experimental setup

\chapter{Experimental Setup}
In this chapter we explain the setup of our experiments. All our experiments are quantitative experiments, measured over a corpus using CloneRefactor. In this chapter, we first explain the corpus we use. We then explain the way in which we calculate the impact of refactorings on the software maintainability.

\section{The corpus}\label{chap:corpus}
For our experiments we use a large corpus of open source projects \cite{githubCorpus2013}. This corpus contains a set of Java projects from GitHub, selected by number of forks. The projects in this corpus were then de-duplicated manually. This results in a variety of Java projects that reflect the quality of average open source Java systems and are useful to perform measurements on.

For our purposes, we have further filtered this corpus to get a controlled set of projects for our experiments. In this section, we explain the steps we took to prepare the corpus\footnote{All scripts to prepare the corpus are available on GitHub: \url{https://github.com/SimonBaars/GitHub-Java-Corpus-Scripts}} and the rationale behind those steps.

\subsection{Filtering Maven projects}
As explained in Section~\ref{chap:challenge}, CloneRefactor requires all the libraries a software project is dependent on in order to perform its clone analysis. As these are not included in the used corpus \cite{githubCorpus2013}, we decided to limit our scope to a single build automation system. We chose Maven, as it is one of the most used build automation systems for Java projects. Maven uses a file, named \texttt{pom.xml}, to describe a projects' dependencies. As no \texttt{pom.xml} files are included in the corpus, we cloned the latest version of each project in the corpus. We then removed each project that has no \texttt{pom.xml} file (which indicates that those do not use the Maven build automation system).

Because we cloned the projects (and do not use the sources provided in the corpus), they still included generated source code. Because we do not want to analyze generated Java files, we decided to further filter the corpus to only include projects with a conventional structure. For Maven, projects are recommended to put their production code in the folder structure \texttt{src/main/java}. We omitted Maven projects that do not adhere to this convention.

\subsection{Gathering Dependencies}
As a next step, we collect all dependencies of the downloaded software projects. Maven uses the following command to automate this process: \texttt{mvn dependency:copy-dependencies -DoutputDirectory=lib}. This process can fail if external dependencies are no longer available. We removed such projects for which we could not obtain all dependencies.

\subsection{Building an AST of all Java files}
To verify the correctness of all Java files in the projects, we used JavaParser \cite{tomassetti2017javaparser} to create an AST of every Java file in the \texttt{src/main/java} folder of the software projects. A very small set of projects had syntactical errors, which we removed from the corpus.

\subsection{Inspecting outliers}
We ran CloneRefactor over every project in the corpus and inspected the output for each project. Some projects gave unusual high numbers of projects of a certain size, which we manually inspected. Often these projects contained generated source files. We removed these projects from the corpus.

\subsection{Resulting corpus}
This procedure results in 2,267 Java projects including all their dependencies. The projects vary in size and quality. The total size of all projects is 14.210.357 lines (11,315,484 when excluding whitespace) over a total of 99,586 Java files. This is an average of 6,268 lines over an average of 44 files per project, 141 lines on average per file. The largest project in the corpus is \textit{VisAD} with 502,052 lines over 1,527.

\section{Minimum clone size}
In this study, we want to find out what thresholds to use to improve maintainability if clones by those thresholds are refactored. However, when clones are very small, they may never be able to improve maintainability. The detrimental effect of the added volume of the newly created method exceeds the positive effect of removing duplication. Because of that, we perform all our experiments with a minimum clone size of 10 tokens, because smaller clones cannot improve maintainability when refactored.

\subsection{Calculating a maintainability score}\label{sec:metricformula}
In this study, we use four metrics to determine maintainability \ref{sec:metrics}. For our experiments, we want to aggregate the scores obtained by these metrics to draw a conclusion about the maintainability increase or decrease after applying a refactoring. We base our aggregation on the following assumptions:
\begin{itemize}
  \item All metrics are equal in terms of weight towards system maintenance effort.
  \item Higher values for the metrics imply lower maintainability.
  \item The obtained increase of the metric divided by the average distance from zero over our corpus weights the metric equally against the other metrics.
\end{itemize}
We calculate the average deviation as follows:
\begin{equation}\label{eq:scoredev}
A = \frac{\sum_{x \in M} abs(x)}{|M|}
\end{equation}
Where \textit{M} is the multiset of all change in a certain metric for all refactorings over all systems in  obtained for a certain metric over our corpus and \textit{|M|} is the cardinality of this set. We then calculate the maintainability for a specific refactoring as follows:
\begin{equation}\label{eq:scoreref}
A_M = \frac{dup}{A_{dup}} + \frac{com}{A_{com}} + \frac{par}{A_{par}} + \frac{vol}{A_{vol}}
\end{equation}
Where \textit{dup} is the decrease in duplication, \textit{com} is the decrease in complexity, \textit{par} is the decrease is method parameters and \textit{vol} is the decrease in system volume after the refactoring is applied. When comparing the maintainability of a set of refactorings we divide the sum of the maintainability score by the amount of clone instances of all refactored clones in the set.