### 2019-08-25 sections/acknowledgements.tex
Acknowledgements

\chapter*{Acknowledgements}
Throughout the writing of this dissertation, I have received a great deal of support and assistance. I would first like to thank my supervisor, Dr. A.M. Oprescu , whose expertise was invaluable in the academic writing and mathematical aspects of this thesis. Thank you for helping me concisely deliver the message I try to deliver.

Next, we would like to thank my peer student Sander Meester for always being there when I was stuck. Whenever I needed to test some ideas, you would always provide valuable insights. Your proofreading efforts helped me a lot to write this thesis.

I would like to thank my supervisor, Xander Schrijen, for all his input during our weekly meetings. We had long discussions about all my ideas, which helped to put my ideas in perspective.

I would like to thank Danny van Bruggen for all his help in my use of JavaParser. I am amazed by his willingness to help others to tackle their problems. JavaParser is at the core of the CloneRefactor tool and without Danny, I would never have come as far in the development of this tool.

Finally, I would like to acknowledge my colleagues from my internship at SIG for their wonderful collaboration. You supported me greatly and were always willing to help me.